---
title: "Vis Creative Assignment 3"
author: "Marissa Getts"
date: "10/3/2020"
output: html_document
---

```{r, include=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(maps)
library(nngeo)
```

### Load Data

```{r}
ccouncils <- st_read("https://gis-slcgov.opendata.arcgis.com/datasets/8a6d4c855ac64dc9a0462b9b1f49bedf_4.kml", 
                  quiet = TRUE) 

parks <- st_read("https://gis-slcgov.opendata.arcgis.com/datasets/04ea8234f39b4d6fac7c1a242d7bbd19_19.kml", 
                 quiet = TRUE)

historic <- st_read("https://gis-slcgov.opendata.arcgis.com/datasets/67ae2a7099e64c6d99cb1f4bb2671850_15.kml", 
                   quiet = TRUE) 

worship <- st_read("https://opendata.arcgis.com/datasets/182293313f774b7492ba892556e31ecf_0.kml?where=CITY%20%3D%20%27SALT%20LAKE%20CITY%27&outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D",
                   quiet = TRUE)
```

### Change Plane

```{r}
UTplane <- "+proj=lcc +lat_1=40.71666666666667 +lat_2=41.78333333333333 +lat_0=40.33333333333334 +lon_0=-111.5 +x_0=500000 +y_0=1000000 +ellps=GRS80 +units=m no_defs"

ccouncils <- ccouncils %>%
  st_transform(UTplane)

parks <- parks %>%
  st_transform(UTplane)

historic <- historic %>%
  st_transform(UTplane)

worship <- worship %>%
  st_transform(UTplane)
```

### Test Map

```{r}
ggplot(ccouncils) +
  geom_sf(fill = NA, color = "light grey") +
  geom_sf(data = parks, fill = "green", color = NA) +
  geom_sf(data = historic, color = "dark grey", size = 0.02) +
  geom_sf(data = worship, color = "magenta", size = 0.02) +
  theme_map() +
  annotation_scale()
```

### 1. Number and Proportion of Historic Sites within 200m of Places of Worship in SLC


```{r}
worship_buffer <- st_buffer(worship, dist = 200) %>%
  st_union()

ggplot(worship_buffer) +
  geom_sf() +
  theme_map()
```
```{r}
worship_historic <- historic[worship_buffer,]
  
ggplot(worship_buffer) +
  geom_sf() +
  geom_sf(data = worship_historic, 
          color = "darkgreen", 
          size = 0.03) +
  theme_map()
```
```{r}
historic <- historic %>%
  st_join(worship_historic) %>%
  mutate(by_worship = !is.na(Name.y))
```

```{r}
n_historicworship <- sum(historic$by_worship)

n_historicworship
```
```{r}
n_historic <- length(historic$by_worship)

pct_historicworship <- n_historicworship / n_historic

pct_historicworship
```



```{r}
left_side  <- st_bbox(historic)$xmin
top_side <- st_bbox(historic)$ymax

ggplot(ccouncils) +
  geom_sf(fill = "white", color = "darkgrey") +
  geom_sf(data = historic, size = 0.2,
          aes(color = by_worship)) +
  scale_color_manual(values = c("magenta", "darkgreen"),
          name = "Historic sites by distance\n to a place of worship", 
          labels = c("no place of worship within 200m",
                     "place of worship within 200m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_historic, big.mark = ","),
                         " historic sites in SLC\n", 
                         prettyNum(n_historicworship, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_historicworship, digits = 0),
                         "%) are within 200\nmeters of a place of worship.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

### 2. Number of Places of Worship by Salt Lake Community Council Districts

```{r}
ccouncils <- ccouncils %>%
  mutate(num_worship = lengths(st_covers(ccouncils, worship)))

ggplot(ccouncils) +
  geom_sf(color = NA, 
          aes(fill = num_worship)) +
  scale_fill_viridis_c(name = "CCs by number of\nplaces of worship",
                       breaks = breaks <- seq(0, 150, by = 15),
                       labels = paste(prettyNum(breaks, big.mark = ",")),
                       option="plasma") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal()) +
  theme_map()

```

### 3. Density of Places of Worship by Community Council Districts in SLC 

```{r}
ccouncils <- ccouncils %>%
  mutate(area = set_units(st_area(ccouncils), km^2)) %>%
  mutate(worship_dens = as.numeric(num_worship / area))

ggplot(ccouncils) +
  geom_sf(color = NA, 
          aes(fill = worship_dens)) +
    scale_fill_viridis_c(name = 
                           "SLC CCs by\nplace of worship density",
                       breaks = breaks <- seq(0, 100, by = 25),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "places of worship per square km"),
                       option="plasma") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) + 
   scale_fill_viridis_c(option = "plasma")

```
```{r}
parks <- parks %>%
  mutate(worship_dist = st_nn(parks, worship, 
                           returnDist = TRUE)$dist) %>%
  mutate(worship_dist = as.numeric(worship_dist))
```
```{r}
avg_park_dist <- mean(parks$worship_dist)

avg_park_dist

park_dist <- parks$worship_dist

park_dist
```


### 4. Average Distance of SLC Parks from a Place of Worship
Honestly - I can't figure how to make the annotation show up for this! I have tried to change the values and can't figure out how to get it to show. 

```{r}
right_side <- st_bbox(parks)$xmax+200
left_side  <- st_bbox(parks)$xmin+200
top_side <- st_bbox(parks)$ymax+200
bottom_side <- st_bbox(parks)$ymin+200

ggplot(ccouncils) +
  geom_sf(fill = NA, color = "darkgrey") +
  geom_sf(data = parks, size = 0.1,
          aes(fill = worship_dist, labels = ))+
  coord_sf(xlim = c(left_side - 2000, right_side -6000), 
           ylim = c(bottom_side - 2000, top_side - 2000), expand = FALSE) +
  scale_x_continuous(fill, limits = c(0,4000, breaks = c(0, 1000, 2000, 3000, 4000))) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text",  vjust = "center", hjust = "center", 
           label = paste("On average, an SLC park\nis ", 
                         prettyNum(avg_park_dist, digits = 3),
                         " meters from a place of worship.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) + 
 scale_fill_continuous(name = "distance to\nplace of\nworship", limits = c(0, 2000), breaks = c(0, 500, 1000, 1500, 2000))
```

### 5.  Number and Proportion of Historic Sites within 200m of Parks in SLC
<br> side note: the number/proportion of this is the same as as with historic sites to places of worship. I have double checked the data and that appears to be correct. An interesting coincidence! 



```{r}
hist_buffer <- st_buffer(historic, dist = 200) %>%
  st_union()

hist_parks <- parks[hist_buffer,]
  
ggplot(hist_parks) +
  geom_sf() +
  geom_sf(data = hist_parks, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()
```
```{r}
historic <- historic %>%
  st_join(hist_parks) %>%
  mutate(by_parks = !is.na(Name.y))
```

```{r}
n_park_hist <- sum(historic$by_parks)

n_park_hist
```
```{r}
n_hist <- length(historic$by_parks)

pct_parkshist <- n_park_hist / n_hist

pct_parkshist
```


```{r}
right_side <- st_bbox(parks)$xmax
left_side  <- st_bbox(parks)$xmin
top_side <- st_bbox(parks)$ymax
bottom_side <- st_bbox(parks)$ymin

ggplot(ccouncils) +
  geom_sf(fill = NA, color = "lightgrey") +
  geom_sf(data = historic, size = 0.2,
          aes(color = by_parks)) +
 coord_sf(xlim = c(left_side - 1500, right_side - 1500), 
           ylim = c(bottom_side - 1500, top_side - 1500), expand = FALSE)+
  scale_color_manual(values = c("lightgreen", "darkgreen"),
          name = "SLC historic sites\nby distance to a parking meter", 
          labels = c("No park within 200 m",
                     "Park within 200 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text", x = left_side + 12000, 
           y = top_side - 8000, 
           label = paste("Of the ", 
                         prettyNum(n_hist, big.mark = ","),
                         " historic sites in SLC\n", 
                         prettyNum(n_park_hist, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_parkshist, digits = 0),
                         "%) are within 200\nmeters of a park.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

### 6. Number of Parks by Community Council District in SLC

```{r}
ccouncils <- ccouncils %>%
  mutate(num_parks = lengths(st_covers(ccouncils, parks)))

ggplot(ccouncils) +
  geom_sf(color = NA, 
          aes(fill = num_parks)) +
  scale_fill_viridis_c(name = "distance to\nplace of\nworship",
                       breaks = breaks <- seq(0, 15, by = 3),
                       labels = paste(prettyNum(breaks, big.mark = ",")),
                       option = "plasma") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 1), 
                                         color = "gray"))
```


